Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-18 ieee_paper.tex
comment
shortening conclusion

In term of memory layout, we adopt a format similar to the ELLPACK
format to store the matrix since the number of nonzeros per row is
constant throughout the matrix. The matrix is given in two arrays. The
first one is the classical {\tt col\_id} array that describes the
sparsity structure of the matrix. Each row of {\tt col\_id} consists of $n_z$
entries that label the column index of each nonzero in $A$. The
other array is the {\tt data} array that gives the values of the 
nonzeros in the four matrices. The values of the matrices are interleaved
so that the value $A^l_{col\_id[k]}$ of the $k$th nonzero of the
$l$th matrix is at index $(k-1)*4+l-1$. The vectors {\tt x} are 
interleaved so that {\tt x}$^l_i$ and {\tt x}$^{l+1}_i$ are
consecutive in memory. The vector {\tt y} is similarly
interleaved. This is schematically shown in Figure~\ref{fig:mat_struct}, 
where the elements of $y$ are grouped in sets of 16, which correspond
to four derivatives applied to four functions, at a single point.

@@

  \caption{Performance of "best code" accross matrix types. The two
    horizontal lines depict worst and best case performance
    (Figure~\ref{fig:gflops-peak-perf}). RCM (blue, $\blacktriangle$)
    and no RCM (red,$\blacktriangledown$).\todo{can we reorder the
      legend to give supercompact first in (a) and compact first in
      (b)?}}
  \label{fig:expe_types}
\end{figure*}

@@

In the previous sections, we have explored in the practical
implementation on a MIC architecture of multiple derivative operators
acting on multiple vectors within the context of RBF-RD. Each
derivative has an associated sparse matrix with a fixed number of
nonzeros on each row. While computing a single derivative of multiple
functions is rather common (vectorization occurs over the vectors), we
accelerate the algorithm further by considering multiple derivatives
with corresponding matrices with identical adjacency graphs. We
specialize to four matrices and four derivatives, with 16 outputs,
computed as a sum of outer products.

Using a matrix with $64^3$ (or $96^3$ rows, we achieve a speedup of
xxx relative to the base case of one derivative and one function (1/1
case), relative to a potential speedup of 16. We have achieved a
speed of 100 Gflop/s on 60 cores using 240 threads, or a speed of 61\% of
maximum possible performance. The maximum possible speed for the 1/1
case is around 40 Gflop/s, of which we achieve 20 Gflop/s. On matrices
that exceed the cache, a reverse Cuthill-McKee is applied to reduce the
bandwidth, with an associated speedup of 30\% up to 135 Gflop/s. Our
optimal implementation makes use of the (as yet unreleased) IMCI MIC instruction set 
%\todo{that's not AVX on MIC, not sure what the name is}, 
which includes a number of swizzling and channel swapping operations, for an
extremely efficient tensor product implementation. A straightforward
implementation without vectorization runs at around 35 Gflops, or 16\% of peak
performance. RCM has no effect because the number of floating point
operations is excessive (the algorithm does not fill the vector
registers to capacity).

In a future work, we will examine the effect of larger stencil sizes,
double precision, and the actual cost of computing the derivatives
within a fluid simulation using RBF-FD.

>>>>

In term of memory layout, we adopt a format similar to the ELLPACK
format to store the matrix since the number of nonzeros per row is
constant throughout the matrix. The matrix is given in two arrays. The
first one is the classical {\tt col\_id} array that describes the
sparsity structure of the matrix. Each row of {\tt col\_id} consists of $n_z$
entries that label the column index of each nonzero in $A$. The
other array is the {\tt data} array that gives the values of the 
nonzeros in the four matrices. The values of the matrices are interleaved
so that the value $A^l_{col\_id[k]}$ of the $k$th nonzero of the
$l$th matrix is at index $(k-1)*4+l-1$. The vectors {\tt x} are 
interleaved so that {\tt x}$^l_i$ and {\tt x}$^{l+1}_i$ are
consecutive in memory. The vector {\tt y} is similarly
interleaved. This is schematically shown in Figure~\ref{fig:mat_struct}, 
where the elements of $y$ are grouped in sets of 16, which correspond
to four derivatives applied to four functions, at a single point. \todo{this description is innacurate.}

@@

  \caption{Performance of "best code" accross matrix types. The two
    horizontal lines depict worst and best case performance
    (Figure~\ref{fig:gflops-peak-perf}). RCM (blue, $\blacktriangle$)
    and no RCM (red,$\blacktriangledown$).\todo{can we reorder the
      legend to give supercompact first in (a) and compact first in
      (b)?}}\todo{is there a symbol/color collision?}
  \label{fig:expe_types}
\end{figure*}

@@

In the previous sections, we have explored in the practical
implementation on a Intel Xeon Phi card of multiple derivative
operators acting on multiple vectors within the context of
RBF-RD. Each derivative has an associated sparse matrix with a fixed
number of nonzeros on each row. While computing a single derivative of
multiple functions is rather common (vectorization occurs over the
vectors), we accelerate the algorithm further by considering multiple
derivatives with corresponding matrices with identical adjacency
graphs. We specialize our study to four matrices and four derivatives,
with 16 outputs, computed as a sum of outer products.

Our implementation makes use of the IMCI MIC instruction set and
includes a number of swizzling and channel swapping operations, for an
extremely efficient tensor product implementation. On a $96^3$ 3D
grid, it reaches a performance of 140 Gflop/s which is 2.7 times
faster than the best possible performance achievable for a single
vector and a single derivative and more than 7 times faster than a
practical implementation.

%% Using a matrix with $96^3$ rows, we achieve a speedup of 7.7 relative
%% to the base case of one derivative and one function (1/1 case). We
%% have achieved a speed between 100 Gflop/s to 140 Gflop/s on 60 cores
%% using 240 threads, or a speed of 47\% to 66\% of the maximum possible
%% performance of this architecture. As a point of comparison, the
%% maximum possible speed for the 1/1 case is around 36 Gflop/s, of which
%% we achieve 18 Gflop/s. Our optimal implementation makes use of the
%% IMCI MIC instruction set which includes a number of swizzling and
%% channel swapping operations, for an extremely efficient tensor product
%% implementation. A straightforward implementation without vectorization
%% runs at around 35 Gflops, or 16\% of peak performance.

In a future work, we will examine the effect of larger stencil sizes
and of using double precision. We will also integrate the techniques
presented in this paper within a fluid simulation using RBF-FD.
